A community-maintained standard library of population genetic models

### 2019-12-19 popSimManu.tex

In Figure \ref{fig:IM_popn_human}C we show estimates of population sizes and divergence
time, for each of the inference methods, using samples drawn from African and European populations
simulated under the \cite{gutenkunst2009inferring} model. Our results highlight many
of the strengths and weaknesses of the different types of methods we used.
For instance, the SFS-based approaches where we fit simple IM models do not capture
recent exponential growth in the CEU population, however do consistently recover the
simulated YRI population size history. Moreover, these approaches allow for estimating
migration rates (Figure \ref{fig:homsap_mig_rates}), also leading to more accurate inference
of divergence times. By contrast, \smcpp is much better at capturing the recent exponential
growth in the CEU population, though the inferred population sizes are in general more
noisy. In addition, the assumption of no migration by \smcpp leads to divergence time
estimates that are consistently underestimated.

Again, we can compare between species and look at the performance of these methods in
the context of a two-population model from \emph{D. melanogaster}. Figure
\ref{fig:two_popn_fly} shows parameter estimates for simulations drawn from
the model of \cite{li2006inferring}, which includes
an ancestral population in Africa, then a population expansion with a population
split and bottleneck into a European population with no post-divergence migration.
Here again, we find that \dadi and \fastsimcoal are more consistent in their inferred
parameters, but ignore the brief population bottleneck in Europe. The inferred
demographic parameters from \smcpp are more noisy, though in some cases are better
able to capture the bottleneck in the Europe population.

@@

From HomSap we ran replicate whole genome simulations under the model
and specified taking 20 samples each from the Europe and Africa populations.
For DroMel, runtimes under OutOfAfrica\_2L06 were prohibitively slow when simulating whole genomes.
For this reason, we chose to present only data from 3 Mb of chromosome 2R for simulations
under OutOfAfrica\_2L06.
For the generic IM simulations, we used the HomSap genome along with the
HapmapII\_GRCh37 genetic map, and sampled 20 individuals from each population.

>>>>

In Figure \ref{fig:IM_popn_human}C we show estimates of population sizes and divergence
time, for each of the inference methods, using samples drawn from African and European populations
simulated under the \cite{gutenkunst2009inferring} model. Our results highlight many
of the strengths and weaknesses of the different types of methods we used.
For instance, the SFS-based approaches where we fit simple IM models do not capture
recent exponential growth in the CEU population, but do consistently recover the
simulated YRI population size history. Moreover, these approaches allow for estimating
migration rates (Figure \ref{fig:homsap_mig_rates}), also leading to more accurate inference
of divergence times. However, these migration rate estimates are somewhat biased likely
due to model misspecification (Figure \ref{fig:homsap_mig_rates}).
By contrast, \smcpp is much better at capturing the recent exponential
growth in the CEU population, though the inferred population sizes are in general more
noisy. In addition, the assumption of no migration by \smcpp leads to divergence time
estimates that are consistently underestimated (Figure \ref{fig:IM_popn_human}C).

Again, we can compare between species and look at the performance of these methods in
the context of a two-population model from \emph{D. melanogaster}. Figure
\ref{fig:two_popn_fly} shows parameter estimates for simulations drawn from
the model of \cite{li2006inferring}, which includes
an ancestral population in Africa, then a population expansion with a population
split and bottleneck into a European population with no post-divergence migration.
Here again, we find that \dadi and \fastsimcoal are more consistent in their inferred
parameters, but ignore the brief population bottleneck in Europe. In addition, \dadi
and \fastsimcoal both do reasonably well at correctly inferring the absence of
migration (Figure \ref{fig:dromel_mig_rates}). By contrast, the inferred
demographic parameters from \smcpp are more noisy, though in some cases better
capture the short bottleneck in the Europe population.

@@

From HomSap we ran replicate whole genome simulations under the OutOfAfrica\_3G09 model
with the HapmapII\_GRCh37 genetic map, taking 20 samples each from the Europe and Africa populations.
For DroMel, runtimes under OutOfAfrica\_2L06 were prohibitively slow when simulating whole genomes
with the Comeron2012\_dm6 map due to the large effective population sizes leading to
high effective recombination rates. For this reason, we chose to present only data
from 3 Mb of chromosome 2R for simulations under OutOfAfrica\_2L06.
For the generic IM simulations, we used the HomSap genome along with the
HapmapII\_GRCh37 genetic map and sampled 20 individuals from each population.
